---
title: "EDS_222_delta-turbidity-analysis"
author: "Steven Cognac"
date: "11/25/2021"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(dataRetrieval) 
library(tidyverse)
library(here)
library(patchwork)
library(readr)
library(feasts)
library(lubridate)
library(tsibble)
library(mapview)
library(sf)

```

USGS API sources
 - https://cran.r-project.org/web/packages/dataRetrieval/vignettes/dataRetrieval.html#daily-data
 - https://dashboard.waterdata.usgs.gov/app/nwd/?region=lower48&aoi=default

## Read in Data

### Water Quality Data
```{r}
wq_data_raw <- read_csv(here("data", "wq_data_raw.csv")) %>% 
  
  # add slough names
  mutate(slough = case_when(site_no == "11455278" ~ "hass",
                            site_no == "11455315" ~ "cache",
                            site_no == "11455276" ~ "shag",
                            site_no == "11455335" ~ "sacramento",
                            site_no == "11455165" ~ "miners",
                            site_no == "11455140" ~ "toe_drain"))
```

https://or.water.usgs.gov/grapher/fnu.html

```{r}
# duplicate turbidity columns. Indicates different project. Need to include all columns that indicate "63680"
# see example here: https://waterdata.usgs.gov/nwis/uv/?site_no=11455315
colnames(wq_data_raw)
```

### Wind Data
```{r}
wind_vaca_raw <- read_csv(here("data", "2805190_vacaville.csv")) %>% 
  select(2,56:58) %>% 
  filter(DATE > "2014-12-31 23:59:00")

wind_sac_raw <- read_csv(here("data", "2805192_sacramento.csv")) %>% 
  select(2,56:58) %>% 
  filter(DATE > "2014-12-31 23:59:00")

wind_concord_raw <- read_csv(here("data", "2805195_concord.csv")) %>% 
  select(2,56:58) %>% 
  filter(DATE > "2014-12-31 23:59:00")

```

## Tidy Dataframes

### Water Quality Data
```{r}
wq_data <- wq_data_raw %>% 
  
  # finds mean across rows with turbidity values.  When only 1 value, that value is used.
  mutate(turbidity_fnu = rowMeans(wq_data_raw[, c("X_63680_00000",
                                                  "X_.BGC.Project._63680_00000",
                                                  "X_BGC.PROJECT_63680_00000",
                                                  "X_CHLOR.INTERCAL.PROJECT...CHLOR.INTERCAL.PROJECT._63680_00000",
                                                  "X_.TS213..YSI.EXO._63680_00000")], na.rm = TRUE)) %>% 
  
  rename("DATE" = "dateTime",
         "gage_height" = "X_00065_00000",
         "water_velocity" = "X_72137_00000",
         "tidal_filt_disch" = "X_72255_00000") %>%
    
  # turbidity readings over 300 FNU are nearly impossible
  filter(turbidity_fnu <= 300,
         DATE > "2014-12-31 23:59:00") %>% 
  
  select("DATE",
         "slough",
         "gage_height",
         "turbidity_fnu",
         # "water_velocity",
         # "tidal_filt_disch"
         )

# convert NaN values to NA
wq_data$turbidity_fnu[is.nan(wq_data$turbidity_fnu)]<-NA

write_csv(wq_data, here("data", "wq_data_cleaned.csv"))

rm(wq_data_raw)
```


```{r Figure1, echo=FALSE, fig.align = 'center', out.width = "30%", fig.cap = "Map of wind and water station locations", results = 'markup'}
# load CSC polygon (created in Google Earth)
cache <- st_read(here("data", "cache.kml"))

# bootstrap geometries
lat <- c(38.34917, 38.29167, 38.30786, 38.29415, 38.24301, 38.25611, 38.513, 38.378, 37.99)
lon <- c(-121.6447, -121.6308, -121.6924, -121.7257, -121.6843, -121.6667, -121.494, -121.962,-122.057)
id <- c("toe_drain", "miner", "shag", "hass", "cache", "sacramento", "Sacramento Execuative AP", "Vacaville Nut Tree AP", "Concord Buchanan AP")
type <- c("water quality","water quality","water quality","water quality","water quality","water quality","wind","wind","wind")

points <- as.data.frame(cbind(id, type,lat, lon))

# combine into single dataframe and convert columns to numeric
points <- points %>% 
  mutate(lat = as.numeric(lat),
         lon = as.numeric(lon)) %>% 
  as_tibble()

# convert to sf object
locations_sf <- st_as_sf(points, coords = c("lon", "lat"), crs = 4326)

# interactive map
my_map <- mapview(locations_sf["type"], layer.name = "Station Type") +
  mapview(cache, layer.name = "Cache Shag Slough Complex")

# map imaage export
mapshot(my_map, file = here("images", "map_plot.jpeg"))

```

```{r}
print(lapply(wq_data, class))
print(unique(wq_data$slough))
```

```{r}
ggplot(wq_data, aes(x = DATE, y = turbidity_fnu, color = slough)) +
  geom_line(alpha = 0.5)
```

## Calcualte averages for turbidity & gage height for all 5 stations
```{r}
wq_data_all <- wq_data %>% 
  group_by(DATE) %>% 
  summarize(turbidity_fnu = mean(turbidity_fnu), 
            gage_height = mean(gage_height),
            slough="all")

# Compare combined statsions to all
# wq_joined <- rbind(wq_data, wq_data_avg) %>% 
#  group_by(slough)

# remove NA values
# wq_joined <- wq_joined[!duplicated(wq_joined$DATE), ]

# create tsibble
# wq_tsbl <- as_tsibble(wq_filt, key = slough, index = DATE)
```



### Wind Data

```{r}
# vacaville nut tree airport
wind_vaca <- wind_vaca_raw %>%
  mutate(station = "vacaville",
         DATE = lubridate::floor_date(DATE, "hours")) %>%
  group_by(DATE, station) %>% 
  summarise(HourlyWindSpeed)

wind_vaca <- wind_vaca[!duplicated(wind_vaca$DATE), ]

# sacramento airport
wind_sac <- wind_sac_raw %>%
  mutate(station = "sacramento",
         DATE = lubridate::floor_date(DATE, "hours")) %>%
  group_by(DATE, station) %>% 
  summarise(HourlyWindSpeed)

# remove duplcate values
wind_sac <- wind_sac[!duplicated(wind_sac$DATE), ]

# concord airport
wind_concord <- wind_concord_raw %>%
  mutate(station = "concord",
         DATE = lubridate::floor_date(DATE, "hours")) %>%
  group_by(DATE, station) %>% 
  summarise(HourlyWindSpeed)

wind_concord <- wind_concord[!duplicated(wind_concord$DATE), ]


# combine dataframes
wind <- rbind(wind_vaca, wind_sac, wind_concord)

write_csv(wind, here("data", "wind_cleaned.csv"))

rm(wind_concord_raw, wind_concord, wind_sac_raw,wind_sac, wind_vaca_raw, wind_vaca)
```


```{r}
class(wind)
lapply(wind, class)
```

## Calcualte average hourly wind speed for three stations
```{r}
wind_all <- wind %>% 
  group_by(DATE) %>% 
  summarize(HourlyWindSpeed = mean(HourlyWindSpeed)) %>% 
  mutate(station = "vaca_sac_concord")

# combine datasets
wind_join <- rbind(wind, wind_all)
wind_tsbl <- wind_join %>% as_tsibble(key = station)

```


## Combine Water Quality and Wind Data

```{r}
df <- left_join(wq_data_all, wind_all, by = "DATE") %>% 
  filter(DATE > "2015-09-01 00:00:00") %>% 
  na.omit()

colnames(df)
```




## Linear regression on entire dataset
```{r}
turbidity_lm1 <- lm(turbidity_fnu ~ HourlyWindSpeed + gage_height, data = df)

turbidity_lm1_sum <- summary(turbidity_lm1)
turbidity_lm1_sum
```


## Scatterplots

Hourly Turbidity scatterplot with regression line.  Shows general downwards trend from 2015-2021
```{r}
# scatter plot with OLS regression line
ols_plot <- ggplot(data = df, aes(x = DATE, y = turbidity_fnu)) +
  geom_line(size = 0.5, alpha = 0.5) +
  geom_smooth(method = lm, formula= y~x, se = FALSE) +
  labs(title = "Turbidity (Hourly)",
       subtitle = "2015-2021",
       y = "Turbidity (FNU)",
       x = "Date")
  

# create predictions and residuals
predictions <- df %>% modelr::add_predictions(turbidity_lm) %>%
  mutate(residuals = turbidity_fnu-pred)

# histogram of residuals
resid_hist <- ggplot(data=predictions) + geom_histogram(aes(residuals), bins=40) +
  labs(subtitle = "Residual Plot")

# qq-plot
qq_plot <- ggplot(predictions, aes(sample = residuals)) +
  geom_qq(color = "deepskyblue3") +
  geom_qq_line() +
    labs(subtitle = "Q-Q Plot",
       x = "Theoretical Quantiles",
       y = "Standardized Residuals")

ols_plot + (resid_hist / qq_plot) + plot_annotation(title = "OLS Regression and Explanatory Plots")
```


# Boxplots of Independent Variables
```{r Figure4, echo=FALSE, fig.align = 'center', out.width = "70%", fig.cap = "Histogram and box plots of independnet and dependent variables", results = 'markup'}
# Histograms
hist_turbidity<- ggplot(data = df, aes(x = turbidity_fnu)) +
  geom_histogram(bins = 40) +
  labs(x = "Turbidity (FNU)")
hist_wind <- ggplot(data = df, aes(x = HourlyWindSpeed)) +
  geom_histogram(bins = 40) +
  labs(x = "Wind Speed (mpg)")
hist_gage <- ggplot(data = df, aes(x = gage_height)) +
  geom_histogram(bins = 40) +
  labs(x = "Gage Height (ft)")

# boxplots
box_turbidity<- ggplot(data = df, aes(x = turbidity_fnu)) +
  geom_boxplot(outlier.colour="red", outlier.shape=3,
                outlier.size=4) +
  labs(x = "Turbidity (FNU)")
box_wind <- ggplot(data = df, aes(x = HourlyWindSpeed)) +
  geom_boxplot(outlier.colour="red", outlier.shape=3,
                outlier.size=4) +
  labs(x = "Wind Speed (mpg)")
box_gage <- ggplot(data = df, aes(x = gage_height)) +
  geom_boxplot(outlier.colour="red", outlier.shape=3,
                outlier.size=4) +
  labs(x = "Gage Height (ft)")


explore_plots <- (hist_turbidity + hist_wind + hist_gage) / 
  (box_turbidity + box_wind + box_gage) + 
  plot_annotation(caption = 'Source: NOAA NCEI and USGS',
                  tag_levels = 'A')
explore_plots
ggsave(here("images","hist_box_plots.png"), width = 7, height = 3)
```


`

```{r}

# time period options: as.Date(.), yearweek(.), yearmonth(.), yearquarter(.), lubridate::floor_date(., "4 hour")

# interval = lubridate::floor_date(., "4 hours")

df_tsbl <- df %>% 
  as_tsibble %>% 
  group_by_key() %>% 
  
  # every 2 hours
  index_by(time = ~as.Date(.)) %>% 
  summarise(turbidity_fnu = mean(turbidity_fnu, na.rm = TRUE),
            gage_height = mean(gage_height, na.rm = TRUE),
            HourlyWindSpeed = mean(HourlyWindSpeed, na.rm = TRUE)) %>% 
  na.omit()

df_tsbl

```
## Plot Combined Datasets
```{r}
hourly_plot <- ggplot(df, aes(x = DATE, y = turbidity_fnu)) +
  geom_line(size = 0.4, alpha = 0.8, show.legend = FALSE) +
  
  geom_line(data = df, aes(x = DATE, y = HourlyWindSpeed, color = "firebrick2"), alpha = 0.9, size = 0.3) +
  geom_line(data = df, aes(x = DATE, y = gage_height, color = "deepskyblue2"), alpha = 0.6) +
  ylim(0, 35) +
  scale_x_datetime(limits = as.POSIXct(c("2019-09-30", "2020-10-01"))) +
  labs(title = "Turbidity, Wind Speed, and Gage Height Comparison",
       subtitle = "Hourly Averages",
       x = "Date",
       y = "Turbidity, Wind, & Gage Height") +
  scale_color_discrete(name = "", labels = c("Gage Height (ft)", "Wind Speed (mph)"))

hourly_plot
```


```{r Figure4, echo=FALSE, fig.align = 'center', out.width = "20%", fig.cap = "Correlation plots of turbidity compared to wind speed and gage height", results = 'markup'}
daily_plot <- ggplot(df_tsbl, aes(x = time, y = turbidity_fnu)) +
  geom_line(size = 0.4, alpha = 0.6) +
  geom_line(data = df_tsbl, aes(x = time, y = HourlyWindSpeed, color = "firebrick2", col="Wind"), alpha = 0.9, size = 0.3) +
  geom_line(data = df_tsbl, aes(x = time, y = gage_height, color = "deepskyblue2"), alpha = 0.6) +
  ylim(0, 35) +
  scale_x_date(limits = as.Date(c("2019-09-30", "2020-10-01"))) +
  labs(subtitle = "Daily Averages",
       x = "Date") +
  ylab(NULL) +
  scale_color_discrete(name = "", labels = c("Gage Height (ft)", "Wind Speed (mph)"))

compare_plot <- (hourly_plot / daily_plot) + 
  plot_layout(guides = "collect")

ggsave(here("images","compare_plot.png"))
compare_plot


```



**2.1a Response**\

 - **The intercept ($beta_0$) estimate tells us that when mean number of Frost days (i.e. days with a minimum temp below freezing) is zero , we predict 11.38 murders per 100,000 population, in 1976.**\
 - **The coefficient($beta_1$) on Frost tells us that for each 1 day increase in mean number of Frost days, we expect the murder rate to decrease by 0.04 murders per 100,000 population, in 1976.**\

## Linear regression on smoothed dataset
```{r}
turbidity_lm <- lm(turbidity_fnu ~ HourlyWindSpeed + gage_height, data = df_tsbl)

turbidity_lm_sum <- summary(turbidity_lm)
turbidity_lm_sum

summ(turbidity_lm)
```


```{r Figure3, echo=FALSE, fig.align = 'center', out.width = "60%", fig.cap = "Correlation plots of turbidity compared to wind speed and gage height", results = 'markup'}
wind_plot_daily <- ggplot(data = df_tsbl, aes(x = HourlyWindSpeed, y = turbidity_fnu, show.legend = FALSE)) +
  geom_point(size = 0.3, alpha = 0.1, show.legend = FALSE) +
  geom_smooth(method = lm, se = FALSE) +
  labs(title = "Wind Speed vs Turbidity",
       subtitle = "Daily",
       x = "Wind Speed (mph)",
       y = "Turbidity (FNU")

gage_plot_daily <- ggplot(data = df_tsbl, aes(x = gage_height, y = turbidity_fnu)) +
  geom_point(size = 0.3, alpha = 0.1, show.legend = FALSE) +
  geom_smooth(method = lm, se = FALSE) +
  labs(title = "Gage Height vs Turbidity",
       subtitle = "Daily",
       x = "Gage Height (feet)",
       y = "Turbidity (FNU)")


wind_plot_hourly <- ggplot(data = df, aes(x = HourlyWindSpeed, y = turbidity_fnu, show.legend = FALSE)) +
  geom_point(size = 0.3, alpha = 0.1, show.legend = FALSE) +
  geom_smooth(method = lm, se = FALSE) +
  labs(subtitle = "Hourly",
       x = "Wind Speed (mph)",
       y = "Turbidity (FNU")

gage_plot_hourly <- ggplot(data = df, aes(x = gage_height, y = turbidity_fnu)) +
  geom_point(size = 0.3, alpha = 0.1, show.legend = FALSE) +
  geom_smooth(method = lm, se = FALSE) +
  labs(subtitle = "Hourly",
       x = "Gage Height (feet)",
       y = "Turbidity (FNU)")

(wind_plot_daily + gage_plot_daily) / (wind_plot_hourly + gage_plot_hourly) +
  theme_minimal()

ggsave(here("images","simple_visual.png"), width = 7, height = 3)

```






