---
title: "data_download"
author: "Steven Cognac"
date: "11/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dataRetrieval) 
library(tidyverse)
library(here)
library(patchwork)
library(readr)
library(janitor)
library(feasts)

```

## Turbidity and Gage Height Data

Acquired from the USGS online portal.  Data aqcuired from January 1, 2015 through September 30, 2021.

```{r}
# Hass, Cache, Shag, Sacramento Deepwater Channel, Miners, and Toe Drain slough in inland delta
siteNumbers <- c(11455278, 11455315, 11455276, 11455335, 11455165, 11455140)

# gage height, turbidity (FNU), mean water velocity (ft per sec) discharge, tidally filtered (cfs)
parameterCd <- c("00065", "63680","72255", "72137") 
startDate <- "2015-10-01" # use "" to grab entire data from beginning of recording
endDate <- "2021-09-30"
tz = "PST"

siteInfo <- readNWISsite(siteNumbers)
siteInfo

```

```{r}
# create database based on above parameters
wq_data_raw <- readNWISuv(siteNumbers, parameterCd, startDate, endDate, tz)
```

```{r}
write_csv(x = wq_data_raw,here("data", "wq_data_raw.csv"))
```

```{r}
colnames(wq_data_raw)
```

## Wind Data

Data downloaded from the NOAA National Centers for Environmental Information (NCEI) Climate Data Online search tool for "Daily Summaries": https://www.ncdc.noaa.gov/cdo-web/search?datasetid=GHCND

Wind data downloaded for three locations
 - Antioch Pumping Plant Number 3 (station ID: GHCND:USW00093241)
 - Sacramento Municipal Airport (station ID: GHCND:USW00023232)
 - Vacaville Nut Tree Airport (Station ID: GHCND:USC00040232)

sacramento airport metadata - https://www.ncdc.noaa.gov/cdo-web/datasets/GHCND/stations/GHCND:USW00023232/detail
# Stations	WBAN:23202
# Begin Date	2015-09-01 00:00
# End Date	2021-09-30 23:59
```{r}
wind_antioch_raw <- read_csv(here("data", "2802840_travis_afb.csv"))

wind_sac_raw <- read_csv(here("data", "2805157_sacramento.csv"))

wind_vaca_raw <- read_csv(here("data", ""))
```

